Paper Analysis

### 2025-04-18 Paper_WriteUp/Paper_Plots.qmd
update code for figure 3

## 1700-1900 Previous Game

@@

```{r figure 3 - Updated}
### Adam
#' April 17, 2025
#' adding code to updated fig3
#' only comparing gamma1, gamma2, & mu_beta across all 4 cohorts

@@

gm_amateur_comp_global <- bind_rows(pars_17_19, pars_20_22, pars_23_25, pars_gm) |> 
  ggplot(aes(x = value, y = name, color = factor(model, levels = c("1700-1900", "2000-2200", "2300-2500", "GM")))) +
  stat_pointinterval(position = position_dodge(width = 0.2)) +
  facet_wrap(~name,
             scales = "free",
             labeller = labeller(name = facet_labels)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_3col() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

gm_amateur_comp_global

```{r figure 3 left - plot global parameters for 17-19}
data_path <- here("results/lichess1700-1900_2025")

bullet_17_19 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))

facet_labels <- as_labeller(c(
  mu_beta = "mu[beta]",
  gamma1 = "gamma[1]",
  gamma2 = "gamma[2]",
  sigma_1 = "sigma[1]",
  `tau[1]` = "tau[1]",
  `tau[2]` = "tau[2]",
  sigma_g1 = "sigma[g[1]]",
  sigma_g2 = "sigma[g[2]]"
), label_parsed)

global_17_19_bullet_prev <- fit_17_19_bullet_prev$draws(c("mu_beta",
                                                          "gamma1", "gamma2",
                           "sigma_1", "tau[1]", "tau[2]",
                           "sigma_g1", "sigma_g2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~name, ncol = 4, scales = "free",
             labeller = labeller(name = facet_labels)) +
  labs(x = element_blank(), y = element_blank()) +
  theme_4col() +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

global_17_19_bullet_prev
ggsave(plot = global_17_19_bullet_prev, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "17_19_prev_global.png")), dpi = 600,
       height = 5, width = 7)
```

```{r interpret the estimated coefficients 17-19}
#| eval: false
gamma_ests <- fit_17_19_bullet_prev$summary(variables = 
                                              c("gamma1",
                                                "gamma2")) |> 
  pull(mean)
## compute change in win probability for gamma1
1/(1 + exp(-1 * (0)))
1/(1 + exp(-1 * (gamma_ests[1])))

@@

@@

## GM previous game

```{r figure 3 right - global parameters for gm data}
data_path <- here("results_2025/lichessGrandmasters")

fit_gm_bullet_prev <- 
  readRDS(file = paste0(data_path,"/all_rated_bullet_model_prev_n1.RDS"))

bullet_gm <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))

global_gm_bullet_prev <- fit_gm_bullet_prev$draws(c("mu_beta",
                                                          "gamma1", "gamma2",
                           "sigma_1", "tau[1]", "tau[2]",
                           "sigma_g1", "sigma_g2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
  ggplot(aes(value)) +
  geom_histogram() +
   facet_wrap(~name, ncol = 4, scales = "free",
             labeller = labeller(name = facet_labels)) +
  labs(x = element_blank(), y = element_blank()) +
  theme_4col() +
  scale_x_continuous(breaks = breaks_pretty(n = 2))

global_gm_bullet_prev
ggsave(plot = global_gm_bullet_prev, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_prev_global.png")), dpi = 600,
       height = 5, width = 7)
```

```{r interpret the estimated coefficients gm}
#| eval: false
gamma_ests_gm <- fit_gm_bullet_prev$summary(variables = c("gamma1",
                                                          "gamma2")) |> 
  pull(mean)
## compute change in win probability for gamma1
1/(1 + exp(-1 * (0)))
1/(1 + exp(-1 * (gamma_ests_gm[1])))
## change in win probability for elo difference
1/(1 + exp(-1 * (0))) ## equally rated, playing as black
1/(1 + exp(-1 * (100 * gamma_ests_gm[2])))
```

>>>>

@@

```{r figure 3}
### Adam
#' April 17, 2025
#' adding code to updated fig3
#' only comparing gamma1, gamma2, & mu_beta across all 4 cohorts

@@

comp_global <- bind_rows(pars_17_19, pars_20_22, pars_23_25, pars_gm) |> 
  ggplot(aes(x = value, y = name, color = factor(model, levels = c("1700-1900", "2000-2200", "2300-2500", "GM")))) +
  stat_pointinterval(position = position_dodge(width = 0.2)) +
  facet_wrap(~name,
             scales = "free",
             labeller = labeller(name = facet_labels)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_3col() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

ggsave(plot = comp_global, 
       filename = paste0(here("Paper_WriteUp", "paper_figures", "fig3.png")), dpi = 600,
       height = 5, width = 7)
```

```{r interpret the estimated coefficients 17-19}
#| eval: false
gamma_ests <- model_17_19$summary(variables = 
                                              c("gamma1",
                                                "gamma2")) |> 
  pull(mean)
## compute change in win probability for gamma1
1/(1 + exp(-1 * (0)))
1/(1 + exp(-1 * (gamma_ests[1])))

@@

```{r interpret the estimated coefficients gm}
#| eval: false
gamma_ests_gm <- model_gm$summary(variables = c("gamma1",
                                                          "gamma2")) |> 
  pull(mean)
## compute change in win probability for gamma1
1/(1 + exp(-1 * (0)))
1/(1 + exp(-1 * (gamma_ests_gm[1])))
## change in win probability for elo difference
1/(1 + exp(-1 * (0))) ## equally rated, playing as black
1/(1 + exp(-1 * (100 * gamma_ests_gm[2])))
```

@@
